---
title: "Homework 1"
author: "Yuqing Liu"
date: 2023-09-14
output: github_document
---

# Problem 1
Load the moderndive library, and use the following code to load the early_january_weather dataset.
```{r}
library(moderndive)
data("early_january_weather")
```
Write a short description of the dataset using inline R code. To do this, I used the str() function to see the structure of the dataset.
```{r}
str(early_january_weather)
summary(early_january_weather)
```

We can find the number of rows (observations) and columns (variables) in the dataset using the nrow() and ncol().
```{r}
nrow(early_january_weather)
ncol(early_january_weather)
```

We can find the mean temperature using the mean().
```{r}
mean(early_january_weather$temp)
```

Discussion:
The important variable in this dataset is temp (temperature). The value of temp ranges from 24.08 to 57.92.
The dataset has 358 rows and 15 columns.
The mean temperature is 39.58.

Make a scatterplot of temp (y) vs time_hour (x); color points using the humid variable (adding color = ... inside of aes in your ggplot code should help). Describe patterns that are apparent in this plot.

load the ggplot2 library:
```{r}
library(ggplot2)
```
create scatterplot:
```{r yx_scatter}
scatterplot <- ggplot(data = early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(x = "Time (hour)", y = "Temperature", color = "Humidity") +
  scale_color_gradient(low = "blue", high = "red")
scatterplot
```
export the scatterplot to the project directory using ggsave:
```{r}
ggsave("scatterplot.png", plot = scatterplot, width = 8, height = 6, units = "in")
```